---
title: "Criminal Urbana (XXI)"
subtitle: "Part 2"
author: ILIA LOMASOV
output:
  ioslides_presentation:
    transition: 0
    # widescreen: true
date: "2023-04-19"
---

<!-- <style> -->
<!-- div.footnotes { -->
<!--   position: absolute; -->
<!--   bottom: 0; -->
<!--   margin-bottom: 10px; -->
<!--   width: 80%; -->
<!--   font-size: 0.6em; -->
<!-- } -->
<!-- </style> -->

<!-- <script src="https://ajax.googleapis.com/ajax/libs/jquery/3.1.1/jquery.min.js"></script> -->
<!-- <script> -->
<!-- $(document).ready(function() { -->
<!--   $('slide:not(.backdrop):not(.title-slide)').append('<div class=\"footnotes\">'); -->

<!--   $('footnote').each(function(index) { -->
<!--     var text  = $(this).html(); -->
<!--     var fnNum = (index+1).toString(); -->
<!--     $(this).html(fnNum.sup()); -->

<!--     var footnote   = fnNum + '. ' + text + '<br/>'; -->
<!--     var oldContent = $(this).parents('slide').children('div.footnotes').html(); -->
<!--     var newContent = oldContent + footnote; -->
<!--     $(this).parents('slide').children('div.footnotes').html(newContent); -->
<!--   }); -->
<!-- }); -->
<!-- </script> -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE
                      # , fig.dim=c(10,4)
                      )
rm(list = ls())

# Visuals, Maps & Dates:
library(stringr); library(data.table); library(ggplot2); library(mapview); library(lubridate)
library(viridis); library(raster); library(sf); library(RColorBrewer); library(dplyr)
library(scales); library(ggpubr); library(cowplot); library(leafsync); library(leaflet.extras2)
library(leaflet); library(wesanderson); library(leafgl); library(mapdeck); library(ggthemes)
library(knitr); library(tictoc)

# Analysis:
# library(class); library(glmnet); library(MASS); library(car); library(caret); library(nnet)

# load("~/Library/Mobile Documents/com~apple~CloudDocs/UIUC/Sem2_sp23/STAT 448/AP02/ap01.RData")

# ------- initially I loaded the RData from AP01. Now for the presentation to compile faster I save the progress from time to time, and
# ------- load whatever version of data is saved at the time. I put the computationally heavy parts as markdown comments, which is why
# ------- in order to reproduce it you would need to un-comment the full chunks of code + some of the code within chunks (i'll mark it by)
# ------- three dashes (---) so that you can see which lines have to be uncommented.

load("~/Library/Mobile Documents/com~apple~CloudDocs/UIUC/Sem2_sp23/STAT 448/AP02/save3.RData")
```

## Previously On Criminal Urbana (XXI)...

<!-- <style type="text/css"> -->
<!-- body p { -->
<!--   color: #000000; -->
<!-- } -->
<!-- </style> -->

<div class="black">

<font size="6">

For Battery, Assault, Robbery, Criminal SA:

- Gave statistical overview, sorted by years & times of day
- Constructed ARMA time series models
- Visualized crimes on maps using `mapview` package in `R`

</font>

</div>

<!-- ## Chosen ARMA Models -->

<!-- - Assault: $X_t = \beta_0+\delta t + \varepsilon_t$ -->

<!-- - Robbery: MA($1^{0.1\%}$) w/ trend$^{1\%}$ $X_t = \beta_0+\delta t + \varepsilon_t$ -->

<!-- - Criminal SA:  -->

<!-- - Battery: ARMA($1^{0.1\%},1^{0.1\%}$), w/o trend -->

<!-- ```{r weekday} -->
<!-- datamap <- data[,c("YEAR.OCCURRED","MONTH.OCCURRED",'DATE.OCCURRED', -->
<!--                    "CRIME.CATEGORY.DESCRIPTION","MAPPING.ADDRESS","TIME.OCCURRED")] -->
<!-- datamap <- datamap[-which(is.na(datamap$MAPPING.ADDRESS)),] -->
<!-- datamap <- datamap[YEAR.OCCURRED>=2000,] -->
<!-- datamap <- datamap[which( -->
<!--   CRIME.CATEGORY.DESCRIPTION=="Sex Offenses"| -->
<!--     CRIME.CATEGORY.DESCRIPTION=="Criminal Sexual Assault"| -->
<!--     CRIME.CATEGORY.DESCRIPTION=="Battery"| -->
<!--     CRIME.CATEGORY.DESCRIPTION=="Assault"| -->
<!--     CRIME.CATEGORY.DESCRIPTION=="Robbery"| -->
<!--     CRIME.CATEGORY.DESCRIPTION=="Homicide" -->
<!--   ),] -->

<!-- datamap <- datamap[str_detect(datamap$MAPPING.ADDRESS, -->
<!--         "\\((.*?)\\)"),] -->

<!-- coord <- str_extract(datamap$MAPPING.ADDRESS, -->
<!--         "\\((.*?)\\)") -->
<!-- coord <- str_remove(coord,"\\(") -->
<!-- coord <- str_remove(coord,"\\)") -->
<!-- coord <- t(as.data.frame(str_split(coord,", "))) -->

<!-- datamap$Latitude  <- as.numeric(coord[,1]) -->
<!-- datamap$Longitude  <- as.numeric(coord[,2]) -->
<!-- rm(coord) -->

<!-- datamap <- datamap[Longitude>(-88.25),] -->
<!-- datamap <- datamap[Longitude<(-88),] -->
<!-- datamap <- datamap[Latitude<40.16,] -->
<!-- datamap <- datamap[Latitude>40.05,] -->

<!-- datamap$TIME.OCCURRED <- str_remove(datamap$TIME.OCCURRED,"\\:00$") -->
<!-- datamap$TIME.OCCURRED <- gsub("\\:","\\.",datamap$TIME.OCCURRED) -->
<!-- datamap$TIME.OCCURRED <- as.numeric(datamap$TIME.OCCURRED) -->

<!-- datesplit <- str_split(datamap$DATE.OCCURRED,"/") -->
<!-- numdate <- vector(length = length(datesplit)) -->
<!-- weekday <- vector(length = length(datesplit)) -->
<!-- for(i in 1:length(datesplit)){ -->
<!--   d <- as.vector(datesplit[[i]]) -->
<!--   weekday[i] <- paste(d[3],d[1],d[2],sep = "-") -->
<!--   numdate[i] <- as.numeric(paste(d[3],d[1],d[2],sep = "")) -->
<!-- } -->

<!-- datamap$weekday <- as.character(wday(weekday, week_start=1)) -->
<!-- datamap$numdate <- as.numeric(numdate) -->
<!-- rm(weekday); rm(numdate) -->


<!-- X <- datamap -->

<!-- X$periods <- as.character(as.numeric(X$TIME.OCCURRED<5 | X$TIME.OCCURRED>=21)+ -->
<!--   2*as.numeric(X$TIME.OCCURRED>=5 & X$TIME.OCCURRED<12)+ -->
<!--   3*as.numeric(X$TIME.OCCURRED>=12 & X$TIME.OCCURRED<17)+ -->
<!--   4*as.numeric(X$TIME.OCCURRED>=17 & X$TIME.OCCURRED<21) -->
<!--   ) -->

<!-- X$periods <- gsub("^1$","Night",X$periods) -->
<!-- X$periods <- gsub("^2$","Morning",X$periods) -->
<!-- X$periods <- gsub("^3$","Noon",X$periods) -->
<!-- X$periods <- gsub("^4$","Evening",X$periods) -->

<!-- datamap <- X -->
<!-- ``` -->

<!-- ## The Ultimate Map | Last 365 Days -->

```{r ultmap setup}
ult <- datamap[which(
 CRIME.CATEGORY.DESCRIPTION=="Battery"
 |CRIME.CATEGORY.DESCRIPTION=="Assault"
 |CRIME.CATEGORY.DESCRIPTION=="Robbery"
  ),]
ult <- ult[numdate>=(max(numdate)-10000),]
ult$periods_dn <- fifelse(ult$periods=="Night", "Night (9:00PM-4:59AM)", "Day (5:00AM-8:59PM)")
# mapview(ult,
#         zcol = "CRIME.CATEGORY.DESCRIPTION",
#         layer.name = "Crimes:",
#         # alpha.regions=0.9,
#         # legend.pos = "bottomleft",
#         # homebutton = FALSE,
#         # col.regions=brewer.pal(5, "RdYlBu"),
#         xcol = "Longitude", ycol = "Latitude", crs = 4296, grid = FALSE)+
# mapview(ult,
#         zcol = "periods_dn",
#         layer.name = "Times of day:",
#         # alpha.regions=0.9,
#         # homebutton = FALSE,
#         # legend.pos = "bottomleft",
#         cex = 5,
#         col.regions=wes_palette(n=2, name="GrandBudapest1", type = "discrete"),
#         xcol = "Longitude", ycol = "Latitude", crs = 4296, grid = FALSE)
```

## Data & Research Goal

<div class="black">

<font size="6">

<ul>
  <li>Crimes - robbery, battery, assault</li>
  <li>Timeframe - last 365 days (11/08/21 - 11/08/22)</li>
  <li>Goal - compare areas by danger:</li>
  <ol>
    <li>Cluster data using k-means$^*$</li>
    <li>Compare crime rates across clusters</li>
  </ol>
</ul>

&nbsp;

<font size="5">
$^*$ - full algorithm in appendix
</font>

</font>

<!-- 3.  \* Include external factors (e.g., city population) -->

<!-- *City population - unchanged according to census.* -->


</div>

<!-- # 1. More Dangerous Areas: Using K-Means To Cluster -->

## Crime Map On A Plane

```{r scaled coordinates on a plane}
X <- as.data.frame(ult[,c('Latitude','Longitude')])
Xmeans <- apply(X[,1:2], 2, mean); Xsds <- apply(X[,1:2], 2, sd)
# for(col in 1:2){X[,col] <- X[,col]-Xmeans[col]; X[,col] <- X[,col]/Xsds[col]}
ggplot(X, aes(x=Latitude, y=Longitude))+geom_point(size=1.5)+
  theme(
    panel.background = element_rect(fill = rgb(red=103,green=102,blue=102, maxColorValue = 105)),
    legend.position="none",
    text = element_text(size = 14),
    plot.title = element_text(size = 17),
    # axis.text.x = element_blank(),
    # scale_x_discrete(position = "top"),
    # axis.ticks.x = element_blank(),
    axis.line = element_line(colour = "black"))+
  ggtitle('Crime Coordinates')
```

## Choosing k:

```{r elbow plot}
maxclust <- 15
wps_list <- matrix(nrow = maxclust, ncol = 2)
for(k in 1:maxclust){
  wps <- vector(mode = "numeric", length = 100)
  for(i in 1:100){set.seed(i); wps[i] <- kmeans(X, k)$tot.withinss}
  wps_list[k,1] <- min(wps); wps_list[k,2] <- which.min(wps)
}

ggplot(as.data.frame(wps_list), aes(y=V1, x=1:maxclust))+geom_line()+
  theme(
    panel.background = element_rect(fill = rgb(red=103,green=102,blue=102, maxColorValue = 105)),
    legend.position="none",
    text = element_text(size = 14),
    plot.title = element_text(size = 17),
    # axis.text.x = element_blank(),
    # scale_x_discrete(position = "top"),
    # axis.ticks.x = element_blank(),
    axis.line = element_line(colour = "black"))+
  scale_x_continuous(breaks=1:maxclust)+
  xlab('Number of Clusters')+
  ylab('Smallest WPS (100 Iterations)')+
  ggtitle('Elbow Plot of K-Means')#+
  # geom_point(aes(x=2, y=wps_list[2,1], color='##FF0000'), size=7.5, shape=1)+
  # geom_point(aes(x=5, y=wps_list[5,1], color='##FF0000'), size=7.5, shape=1)
```

```{r map setup, include=FALSE}
# # sharpest/biggest elbow
# k <- which(diff(diff(wps_list[,1]))==sort(diff(diff(wps_list[,1])),decreasing = TRUE)[1])+1
# # 2nd sharpest/biggest elbow
# k <- which(diff(diff(wps_list[,1]))==sort(diff(diff(wps_list[,1])),decreasing = TRUE)[2])+1
maplist <- list("1","2","3","4","5")

k <- 2
set.seed(wps_list[k-1,2]); ult$cluster <- as.character(kmeans(X, k)$cluster)
maplist[[k]] <- mapview(ult,
        zcol = "cluster",
        layer.name = paste("Clusters (out of ", k,"):", sep = ""),
        # alpha.regions=0.9,
        # legend.pos = "bottomleft",
        # homebutton = FALSE,
        col.regions=brewer.pal(3, "Paired")[c(1,3)],
        xcol = "Longitude", ycol = "Latitude", crs = 4296, grid = FALSE)

for(k in c(4,5)){
set.seed(wps_list[k-1,2]); ult$cluster <- as.character(kmeans(X, k)$cluster)
# ult$cluster <- as.character(dbscan(X, eps = 0.25, MinPts = 10)$cluster)

maplist[[k]] <- mapview(ult,
        zcol = "cluster",
        layer.name = paste("Clusters (out of ", k,"):", sep = ""),
        # alpha.regions=0.9,
        # legend.pos = "bottomleft",
        # homebutton = FALSE,
        col.regions=brewer.pal(k, "Paired"),
        xcol = "Longitude", ycol = "Latitude", crs = 4296, grid = FALSE)
}
```

## Clustering with k = 2

```{r map after clustering k 2}
maplist[[2]]
```

## Clustering with k = 4

```{r map after clustering k 4}
maplist[[4]]
```

## Clustering with k = 5

```{r map after clustering k 5}
maplist[[5]]
```

## Testing (k=2)

<div class="black">

```{r dangerous areas table k 2}
k <- 2; set.seed(wps_list[k-1,2]); ult$cluster <- as.character(kmeans(X, k)$cluster)
crimes_freq <- as.data.frame(table(ult$cluster))
crimes_freq <- t(crimes_freq)
percentage <- paste(round(100*as.numeric(crimes_freq[2,])/nrow(ult)), '%', sep='')
crimes_freq <- rbind(crimes_freq, percentage)
rownames(crimes_freq) <- c('Area','Count', 'Percentage')
kable(crimes_freq, row.names = TRUE, col.names = NULL)
```

```{r prop test k 2, results='hide'}
prop.test(x=as.numeric(crimes_freq[2,1]), n=nrow(ult), p=0.5, alternative = "less")
```

&nbsp;

- Z-test for $H_0: p_1=0.5$ is rejected at $1\%$ in favor of $H_1: p_1<0.5$

</div>

## Testing & Conclusions (k=5)

<div class="black">

```{r dangerous areas table k 5}
k <- 5; set.seed(wps_list[k-1,2]); ult$cluster <- as.character(kmeans(X, k)$cluster)
crimes_freq <- as.data.frame(table(ult$cluster))
crimes_freq <- t(crimes_freq)
percentage <- paste(round(100*as.numeric(crimes_freq[2,])/nrow(ult)), '%', sep='')
crimes_freq <- rbind(crimes_freq, percentage)
rownames(crimes_freq) <- c('Area','Count', 'Percentage')
kable(crimes_freq, row.names = TRUE, col.names = NULL)
```

```{r chisq test k 5, results='hide'}
chisq.test(as.numeric(crimes_freq[2,]), p = sample(1/k, k, replace = TRUE))
```

&nbsp;

- $\chi^2$ GoF test for equal proportions - rejected at $0.1\%$.

- **Conclusion:** some areas are definitely more dangerous
- **Downside:** uninhabited areas

</div>

<!-- ## Total, Night & Day: Comparison -->

<!-- <div class="black"> -->

<!-- ```{r crimes by area night} -->
<!-- X1 <- ult -->
<!-- ult <- ult[periods_dn!="Day (5:00AM-8:59PM)",] -->
<!-- crimes_freq <- as.data.frame(table(ult$cluster)) -->
<!-- crimes_freq <- t(crimes_freq) -->
<!-- percentage_night <- paste(round(100*as.numeric(crimes_freq[2,])/nrow(ult)), '%', sep='') -->
<!-- ult <- X1 -->
<!-- ult <- ult[periods_dn=="Day (5:00AM-8:59PM)",] -->
<!-- crimes_freq <- as.data.frame(table(ult$cluster)) -->
<!-- crimes_freq <- t(crimes_freq) -->
<!-- percentage_day <- paste(round(100*as.numeric(crimes_freq[2,])/nrow(ult)), '%', sep='') -->
<!-- ult <- X1 -->

<!-- percentage_total <- rbind(1:5, percentage, percentage_day, percentage_night) -->
<!-- rownames(percentage_total) <- c('Area', 'All', 'Day (5AM-9PM)', 'Night') -->

<!-- kable(percentage_total, row.names = TRUE, col.names = NULL) -->
<!-- ``` -->

<!-- </div> -->

# Thank you!

## Appendix

<div class="black">

<font size="6">

<ol>
  <li>For $k\in (1,15)$</li>
    <ol type='i'>
      <li>Cluster the data using k-means and 100 unique RS</li>
      <li>Select $RS_k=\text{argmin}_{s\in(1,100)}WPS(k,s)$</li>
    </ol>
  <li>Plot $WPS(k,RS_k)$ against $k$, use elbow method to select $k^*$</li>
  <!-- <li>Use elbow method to select $k^*$</li> -->
  <li>Cluster data using $k^*$-means</li>
  <li>Test if clusters have equal number of crimes</li>
</ol>

</font>

</div>

<!-- # 2. Connecting Time and Location -->

<!-- ## Setup -->

<!-- <div class="black"> -->

<!-- - Classification task: use coordinates to predict times of day of crimes -->
<!-- - Supervised learning algorithms: KNN, Logistic Model (w/ and w/o intercept) -->
<!-- - Diagnostics goal: find the model that produces the smallest prediction errors -->

<!-- </div> -->

<!-- ## Results -->

<!-- ```{r class data} -->
<!-- ult <- datamap[which( -->
<!--  CRIME.CATEGORY.DESCRIPTION=="Battery" -->
<!--  |CRIME.CATEGORY.DESCRIPTION=="Assault" -->
<!--  |CRIME.CATEGORY.DESCRIPTION=="Robbery" -->
<!--  |CRIME.CATEGORY.DESCRIPTION=="Criminal Sexual Assault" -->
<!--   ),] -->
<!-- # ult <- ult[numdate>=(max(numdate)-10000),] -->
<!-- cl <- ult$periods -->
<!-- ``` -->

<!-- ```{r modelling all variables, include=FALSE} -->
<!-- tic() -->

<!-- acc <- as.data.frame(matrix(nrow = 4, ncol = 3)) -->
<!-- knnmax <- vector(length = 8) -->

<!-- for(j in 1:2){ -->
<!-- i <- 1 -->
<!-- for(crime in c("Battery","Assault","Robbery","Criminal Sexual Assault")){ -->
<!--   ult <- datamap[which(CRIME.CATEGORY.DESCRIPTION==crime),]; cl <- ult$periods -->

<!--   X <- as.data.frame(ult[,c('Latitude', 'Longitude', 'MONTH.OCCURRED', 'weekday')]) -->
<!--   X$MONTH.OCCURRED <- as.character(X$MONTH.OCCURRED) -->

<!--   Xmeans <- apply(X[,1:2], 2, mean); Xsds <- apply(X[,1:2], 2, sd) -->
<!--   for(col in 1:2){X[,col] <- (X[,col]-Xmeans[col])/Xsds[col]} -->

<!--   # j=1 => time-sequential train/test sets; j=2 => random-sample train/test sets -->
<!--   if(j==1){tr_s <- as.integer(1:(round(0.8*nrow(X))))} -->
<!--   if(j==2){set.seed(100); tr_s <- sample(1:nrow(X), round(0.8*nrow(X)), replace = FALSE)} -->

<!--   acc_knn <- vector(length = 100) -->
<!--   for(k in 1:100){ -->
<!--   mod <- knn(train = X[tr_s,], test = X[-tr_s,], cl = cl[tr_s], k=k) -->
<!--   acc_knn[k] <- sum(mod==cl[-tr_s])/length(cl[-tr_s]) -->
<!--   } -->

<!--   knnmax[i+4*(j-1)] <- which.max(acc_knn) -->
<!--   z_knn <- knn(train = X[tr_s,], test = X[-tr_s,], cl = cl[tr_s], k=knnmax[i+4*(j-1)]) -->

<!--   for(col in 1:2){X[,col] <- X[,col]*Xsds[col]+Xmeans[col]} -->
<!--   dtr <- X[tr_s,]; dtr$cl <- cl[tr_s] -->
<!--   dts <- X[-tr_s,] -->

<!--   z_m <- multinom(cl ~ ., dtr, maxit=100, MaxNWts=8000) -->
<!--   z_m_i <- multinom(cl ~ .-1, dtr, maxit=100, MaxNWts=8000) -->

<!--   acc[i+4*(j-1),1] <- sum(predict(z_m_i, newdata = dts)==cl[-tr_s])/length(cl[-tr_s]) -->
<!--   acc[i+4*(j-1),2] <- sum(predict(z_m, newdata = dts)==cl[-tr_s])/length(cl[-tr_s]) -->
<!--   acc[i+4*(j-1),3] <- sum(z_knn==cl[-tr_s])/length(cl[-tr_s]) -->
<!--   i <- i+1 -->
<!-- } -->
<!-- } -->

<!-- toc() -->

<!-- knnmax_all <- knnmax # 72 28 38 29 93 96 35 29 -->
<!-- ``` -->

<!-- ```{r individual crimes modelling minus month, include=FALSE} -->
<!-- tic() -->

<!-- acc <- as.data.frame(matrix(nrow = 4, ncol = 3)) -->
<!-- knnmax <- vector(length = 8) -->

<!-- for(j in 1:2){ -->
<!-- i <- 1 -->
<!-- for(crime in c("Battery","Assault","Robbery","Criminal Sexual Assault")){ -->
<!--   ult <- datamap[which(CRIME.CATEGORY.DESCRIPTION==crime),]; cl <- ult$periods -->

<!--   X <- as.data.frame(ult[,c('Latitude', 'Longitude', 'weekday')]) -->

<!--   Xmeans <- apply(X[,1:2], 2, mean); Xsds <- apply(X[,1:2], 2, sd) -->
<!--   for(col in 1:2){X[,col] <- (X[,col]-Xmeans[col])/Xsds[col]} -->

<!--   # j=1 => time-sequential train/test sets; j=2 => random-sample train/test sets -->
<!--   if(j==1){tr_s <- as.integer(1:(round(0.8*nrow(X))))} -->
<!--   if(j==2){set.seed(100); tr_s <- sample(1:nrow(X), round(0.8*nrow(X)), replace = FALSE)} -->

<!--   acc_knn <- vector(length = 100) -->
<!--   for(k in 1:100){ -->
<!--   mod <- knn(train = X[tr_s,], test = X[-tr_s,], cl = cl[tr_s], k=k) -->
<!--   acc_knn[k] <- sum(mod==cl[-tr_s])/length(cl[-tr_s]) -->
<!--   } -->

<!--   knnmax[i+4*(j-1)] <- which.max(acc_knn) -->
<!--   z_knn <- knn(train = X[tr_s,], test = X[-tr_s,], cl = cl[tr_s], k=knnmax[i+4*(j-1)]) -->

<!--   for(col in 1:2){X[,col] <- X[,col]*Xsds[col]+Xmeans[col]} -->
<!--   dtr <- X[tr_s,]; dtr$cl <- cl[tr_s] -->
<!--   dts <- X[-tr_s,] -->

<!--   z_m <- multinom(cl ~ ., dtr, maxit=100, MaxNWts=8000) -->
<!--   z_m_i <- multinom(cl ~ .-1, dtr, maxit=100, MaxNWts=8000) -->

<!--   acc[i+4*(j-1),1] <- sum(predict(z_m_i, newdata = dts)==cl[-tr_s])/length(cl[-tr_s]) -->
<!--   acc[i+4*(j-1),2] <- sum(predict(z_m, newdata = dts)==cl[-tr_s])/length(cl[-tr_s]) -->
<!--   acc[i+4*(j-1),3] <- sum(z_knn==cl[-tr_s])/length(cl[-tr_s]) -->
<!--   i <- i+1 -->
<!-- } -->
<!-- } -->

<!-- toc() -->

<!-- knnmax_nomonth <- knnmax -->
<!-- ``` -->


<!-- ```{r individual crimes modelling minus weekday, include=FALSE} -->
<!-- # tic() -->

<!-- acc <- as.data.frame(cbind(matrix(nrow = 4, ncol = 3),acc)) -->
<!-- # knnmax <- vector(length = 4) -->
<!-- knnmax <- c(73,89,18,16) -->

<!-- i <- 1 -->
<!-- for(crime in c("Battery","Assault","Robbery","Criminal Sexual Assault")){ -->
<!--   ult <- datamap[which(CRIME.CATEGORY.DESCRIPTION==crime),]; cl <- ult$periods -->

<!--   X <- as.data.frame(ult[,c('Latitude', 'Longitude', 'MONTH.OCCURRED')]) -->
<!--   X$MONTH.OCCURRED <- as.character(X$MONTH.OCCURRED) -->

<!--   set.seed(100); tr_s <- sample(1:nrow(X), round(0.8*nrow(X)), replace = FALSE) -->

<!--   Xmeans <- apply(X[,1:2], 2, mean); Xsds <- apply(X[,1:2], 2, sd) -->
<!--   for(col in 1:2){X[,col] <- (X[,col]-Xmeans[col])/Xsds[col]} -->

<!--   # acc_knn <- vector(length = 100) -->
<!--   # for(k in 1:100){ -->
<!--   # mod <- knn(train = X[tr_s,], test = X[-tr_s,], cl = cl[tr_s], k=k) -->
<!--   # acc_knn[k] <- sum(mod==cl[-tr_s])/length(cl[-tr_s]) -->
<!--   # } -->

<!--   # knnmax[i] <- which.max(acc_knn) -->
<!--   z_knn <- knn(train = X[tr_s,], test = X[-tr_s,], cl = cl[tr_s], k=knnmax[i]) -->

<!--   for(col in 1:2){X[,col] <- X[,col]*Xsds[col]+Xmeans[col]} -->
<!--   dtr <- X[tr_s,]; dtr$cl <- cl[tr_s] -->
<!--   dts <- X[-tr_s,] -->

<!--   z_m <- multinom(cl ~ ., dtr, maxit=100, MaxNWts=8000) -->
<!--   z_m_i <- multinom(cl ~ .-1, dtr, maxit=100, MaxNWts=8000) -->

<!--   acc[i,1] <- sum(predict(z_m_i, newdata = dts)==cl[-tr_s])/length(cl[-tr_s]) -->
<!--   acc[i,2] <- sum(predict(z_m, newdata = dts)==cl[-tr_s])/length(cl[-tr_s]) -->
<!--   acc[i,3] <- sum(z_knn==cl[-tr_s])/length(cl[-tr_s]) -->
<!--   i <- i+1 -->
<!-- } -->

<!-- # toc() -->
<!-- ``` -->

<!-- ```{r accuracy comparison, results='hide'} -->
<!-- paste(apply(acc, 1, which.max),": ", apply(acc, 1, max), sep = '') -->
<!-- ``` -->

<!-- <div class="black"> -->

<!-- - The models that ended up being chosen: -->

<!-- |Crime|Chosen Model|Accuracy| -->
<!-- |--|--|--| -->
<!-- |Battery|KNN with $k=44$, `MONTH.OCCURRED` not included|$43.49\%$| -->
<!-- |Assault|KNN with $k=43$, `MONTH.OCCURRED` not included|$39.72\%$| -->
<!-- |Robbery|Logit w/o intercept, all variables included|$48.74\%$| -->
<!-- |Criminal SA|KNN with $k=44$, w/o `MONTH.OCCURRED`|$50.78\%$| -->

<!-- </div> -->
